### 2019-06-12 2-Chapters/3-Chapter/chapter3.tex
Multi-player -> multi-players

We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
MAB problems (single- or multi-players),
and reward distributions.
%
We then detail how it implements MAB algorithms (\eg, \UCB),
and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
%
%
The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-player) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.

\subsection{Single- and multi-player MAB problems}

@@

\textbf{Simulation loop for multi-player MAB.}
%
For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
%
Different collision models have been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
We refer to Chapter~\ref{chapter:5} which introduces the multi-player bandit model.

@@

Any simulation library targeting multi-player bandit problems must implement at least another component:
% a collision model, and
a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    % \item
    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask each player $\cA^{(j)}$ her chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment. The
    sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model. Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,
    % \item
    At the end, compute the accumulated reward, the regret, plot visualizations etc.
% \end{itemize}

@@

\textbf{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
Travis CI allows to automatically run short numerical simulations after every commit.
This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-player model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

    \item
In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-player bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}. We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).

@@

$(i)$
Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-player multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-player with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-player case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

@@

\paragraph{Future works.}
%
An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-player or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-player bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-player bandit model,
and even if some have already been implemented, a major future work is to implement the most interesting ones
(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).\\
%
\indent
Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

>>>>

We start by explaining how SMPyBandits simulates MAB problems, by detailing its components:
MAB problems (single- or multi-players),
and reward distributions.
%
We then detail how it implements MAB algorithms (\eg, \UCB),
and what kind of information is displayed, saved and plotted after a batch of simulations of bandit problems.
%
%
The main goal of the library is to be easily able to simulate MAB algorithms (\eg, three algorithms like \UCB, \klUCB{} and Thompson sampling) on one or more bandit models (single- or multi-players) defined by the number of arms $K$ and the distributions $\nu_k$, for some time from $t=1$ to the horizon $T$.
After the simulation, the library then displays statistical summary of the (mean) rewards accumulated by each algorithm, as well as regret and other visualizations.
The same problem is simulated for $N$ independent repetitions (\eg, $N=100$) in order to show mean results with a low variance.

\subsection{Single- and multi-players MAB problems}

@@

\textbf{Simulation loop for multi-players MAB.}
%
For Cognitive Radio dealing with OSA problems and other applications, a well-studied extension is to consider $M\geq2$ players, interacting on the same $K$ arms.
Whenever two or more players select the same arm at the same time (\eg, for OSA, the same frequency channel), they all suffer from a radio collision.
%
Different collision models have been proposed, and the simplest one consists in giving a $0$ reward to each colliding players.
Without any centralized supervision or coordination between players, they must learn to access the $M$ best resources (\ie, arms with highest means) without collisions.
We refer to Chapter~\ref{chapter:5} which introduces the multi-players bandit model.

@@

Any simulation library targeting multi-players bandit problems must implement at least another component:
% a collision model, and
a simulation loop that essentially looks like this.
% \begin{itemize}
    % \item
    First, initialize the MAB problem with $M$ players, and one or more cohorts of $M$ players (one player is one algorithm, usually $M$ times the same one),
    % \item
    Then, for $t=1$ to $t=T$ (known before hand), repeat the following block (for each algorithm). Ask each player $\cA^{(j)}$ her chosen arm $A^{(j)}(t)$, then query the collision model\footnote{~The default collision model is the most widely studied in the literature, where a player encounters a collision (\ie, received a zero reward $Y^{(j)}(t)=0$) if she is not the only one to have chosen an arm $k=A^{(j)}(t)$, otherwise $Y^{(j)}(t)=r_{A^{(j)}(t)}(t)$ if she is the only one playing this arm.} to know which player will get a zero reward (for a collision) and which player will get a random reward from the environment. The
    sample (random) feedback $Y_{k,t}$ (\iid) from distributions $\nu_{k}$, and compute the rewards $r^{(j)}(t)$ from the random feedback and the collision model. Finally feeds the observation $(A^{(j)}(t), Y_{A^{(j)}(t)}(t), r^{(j)}(t))$ to player $j$, for all the $M$ players,
    % \item
    At the end, compute the accumulated reward, the regret, plot visualizations etc.
% \end{itemize}

@@

\textbf{Travis CI.}
Since $2018$, we use the free web service provided by Travis CI (\href{https://travis-ci.org/}{\texttt{Travis-CI.org}}).
Travis CI allows to automatically run short numerical simulations after every commit.
This allows to check that each modification on any part of the codebase does not break anything, as well as giving an up-to-date example of log files that shows online the results of different examples of experiments. The tests covers all the main models and almost all the algorithms implemented in SMPyBandits, and it has been proven very useful to quickly find and fix new bugs.
For more details, see \href{https://travis-ci.org/SMPyBandits/SMPyBandits}{\texttt{Travis-CI.org/SMPyBandits/SMPyBandits}}.
%
Since its first use, the service did about 160 builds, and about $30\%$ of them were useful to detect newly introduced issues or bugs in the code.
The build is configured using the \texttt{.travis.yml} file in SMPyBandits main folder, and it runs numerical experiments with a short horizon (\eg, $T=100$ or $T=1000$) and a small number of repetitions (\ie, $N=4$), but covering all the different models implemented by SMPyBandits (and even including some that are not covered in this chapter, like the Markov model, or the sparse multi-players model).
Builds typically run for $15$ minutes, and Travis CI have been proven to be very useful in our development process.

@@

    \item
In \cite{Besson2018ALT} and in Chapter~\ref{chapter:5} below, we used SMPyBandits for all the simulations for multi-players bandit algorithms\footnote{~Similarly, see the page \texttt{\href{https://SMPyBandits.GitHub.io/MultiPlayers.html}{SMPyBandits.GitHub.io/MultiPlayers.html}} on the documentation.}. We designed the two \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.RandTopM.html}{RandTopM}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.MCTopM.html}{MCTopM}} algorithms and proved than they obtain logarithmic regret in the usual setting, and outperform significantly the previous state-of-the-art solutions (\ie, \texttt{\href{https://SMPyBandits.GitHub.io/docs/PoliciesMultiPlayers.rhoRand.html}{rhoRand}}, \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MEGA.html}{MEGA}} and \texttt{\href{https://SMPyBandits.GitHub.io/docs/Policies.MusicalChair.html}{MusicalChair}}).

@@

$(i)$
Its purpose is to easily implement numerical simulations of stochastic or piece-wise stochastic problems of single- or multi-players multi-armed bandits.
SMPyBandits is distributed on GitHub and Pypi freely, under an open-source license, and it is extensively documented (at \href{https://SMPyBandits.GitHub.io}{\texttt{SMPyBandits.GitHub.io}}).
Our library allows any researcher to easily run numerical simulations of different kinds of multi-armed bandit problems, requiring only a small knowledge of Python thanks to its documentation, its well-designed API, and many examples of simulation scripts and configuration files included.

$(ii)$
We detailed how SMPyBandits is implementing arms, problems, algorithms, and use these components to implement a simulation loop, with various visualizations being performed after the simulation.
As far as now, SMPyBandits is restricted to the finite-arm cases, but it supports a wide range of arms distributions.
Different kinds of models are implemented, from stationary single-player to piece-wise stationary multi-players with different collision models.
One of the main qualities of the library is that it is quite exhaustive, as all the main families of algorithms covering these different models have been implemented, even very recent algorithms from the literature, as we tried our best in following the active research since December $2016$ to June $2019$.
More than $65$ algorithms or variants of algorithms are implemented for the single-player case, $5$ for the expert aggregation problem, about $15$ for the multi-players case, and about $20$ for the piece-wise stationary problem.
All the codebase is fully documented, and the library is using continuous integration to run automated tests on the code after every modification.
%
When comparing algorithms on a problem, the main performance measure is the regret, but the library also computes, stores and visualizes other measures, such as best-arm selection rate or mean cumulated reward, as well as real time and memory costs.

@@

\paragraph{Future works.}
%
An interesting future work left on our library SMPyBandits could be to implement variants of the single-player stochastic models, as well as variants for the multi-players or the non-stationary cases.
For more details, see the issue tickets at \href{https://github.com/SMPyBandits/SMPyBandits/issues/}{\texttt{GitHub.com/SMPyBandits /SMPyBandits/issues/}}.
For instance it would be interesting to extend the library for problems with non-finite arms, \eg, linear or contextual bandit (\href{https://github.com/SMPyBandits/SMPyBandits/issues/117}{ticket 117}),
or to add support for the ``dynamic case'' of multi-players bandits to allow arrivals or departures of players (\href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{ticket 124}).
%
At the end of Chapter~\ref{chapter:5}, we present many extensions of the multi-players bandit model,
and even if some have already been implemented, a major future work is to implement the most interesting ones
(tickets \href{https://github.com/SMPyBandits/SMPyBandits/issues/120}{120}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/124}{124}, \href{https://github.com/SMPyBandits/SMPyBandits/issues/185}{185}).\\
%
\indent
Since $2017$, I also received by emails or by GitHub issues some requests to implement new features or models. Most of them were implemented quickly, but some would require more time and are left as possible future works.
Two example of requests include a model where the number of arm can change at some or all time step (an example of such model is called ``sleeping bandit'', \href{https://github.com/SMPyBandits/SMPyBandits/issues/123}{issue 123}), to implement the model of rotting bandits (\href{https://github.com/SMPyBandits/SMPyBandits/issues/156}{issue 156}), or to add a web-based user interface for an even easier usage of the library requiring no command line or knowledge of Python (\href{https://github.com/SMPyBandits/SMPyBandits/issues/133}{issue 133}).

### 2019-06-12 2-Chapters/4-Chapter/CrownCom_17.git/IoT_slotted.tex
Multi-player -> multi-players

\textbf{Some assumptions on the network occupation:}
%
We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (for instance, about $1000$ to $100000$, while $K$ is about $4$ to $256$).
As this problem is only interesting if devices are able to communicate reasonably efficiently with the gateway, we assume that devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any device using the \SI{868}{\mega\hertz} ISM band in Europe (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
%
We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for realistic IoT networks.\\
% \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
\indente
On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$). Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $D \leq K$ dynamic devices.} than the number of channels}.

>>>>

\textbf{Some assumptions on the network occupation:}
%
We focus on \emph{dense networks}, in which the number of devices $S + D$ is very large compared to $K$ (for instance, about $1000$ to $100000$, while $K$ is about $4$ to $256$).
As this problem is only interesting if devices are able to communicate reasonably efficiently with the gateway, we assume that devices only communicate occasionally, \ie, with a low \emph{duty cycle}, as it is always considered for IoT.
Note that even unlicensed bands have such limitations, as for instance this is a strict requirement for any device using the \SI{868}{\mega\hertz} ISM band in Europe (and its regulation also enforces a maximum transmission power, but we do not consider power transmission in this work).
%
We prefer this choice rather than non-crowded networks, \ie, where $S + D \leq K$, as the former makes more sense for realistic IoT networks.\\
% \footnote{~Here, $K$ will typically be about $10$ different radio channels, usually in the same RF band and with separate mean carrier frequencies, and the number of devices will be of the order of $1000$ to $10000$.}.
\indent
On the opposite, imagine if there were only $K=4$ channels being occupied by $D+S = 100$ devices, each communicating with a high rate of $p=1/10$, with a non-zero occupancy in each channel (\ie, $\forall k, S_k > 0$). Then, almost all time slots will lead to collisions in all channels, for a uniformly random access scheme, and thus the network efficiency (\ie, successful communication rate) will be so close to zero that using learning algorithms cannot improve much.
Such scenario is not our target of interest, and thus we prefer to only consider feasible scenarios where $p$ is smaller than $K/(S+D)$, in order to \emph{have an average of active devices not larger\footnote{~In Chapter~\ref{chapter:5} we simply consider the case of $p=1$ and $D \leq K$ dynamic devices.} than the number of channels}.

### 2019-06-12 2-Chapters/4-Chapter/chapter4.tex
Commentaire Christophe Ch4.1 et 4.2
Multi-player -> multi-players

\textbf{Publications.}
%
This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC}.

@@

\paragraph{Multi-Player MAB with \emph{collision avoidance}.}

We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-player MAB} model to describe our problem, as proposed by \cite{Zhao10}.

The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
and their random activation patter makes the number of active devices a random variable.
Analyzing multi-player MAB models under this hypothesis is much harder, and is left as a future work.
%
We study this other direction in the next Chapter~\ref{chapter:5}.

>>>>

\textbf{Publications.}
%
This chapter is mainly based on our articles \cite{Bonnefoi17,Besson2018ICT,Besson2019WCNC,Bonnefoi2019WCNC,MoyBesson2019}.

@@

\paragraph{Multi-Players MAB with \emph{collision avoidance}.}

We are actually able to analyze a simpler model, if we assume to have at most $K$ devices with a transmission probability of $p=1$.
Instead of the direction pursued in this chapter, another possibility is to consider a \emph{multi-players MAB} model to describe our problem, as proposed by \cite{Zhao10}.

The main difference between the two models is the fact that in this chapter, devices do not transmit their messages at every time step, but following a random activation process (with a fixed transmission probability $p < 1$).
If static and dynamic devices that have to transmit at a fixed time are denoted \emph{active devices},
and their random activation patter makes the number of active devices a random variable.
Analyzing multi-players MAB models under this hypothesis is much harder, and is left as a future work.
%
We study this other direction in the next Chapter~\ref{chapter:5}.
